% paper2_gravity_tau.tex
% Information Loss is Local:
% The Gravitational Refractive Index as Petz Recovery Fidelity
% Paper II in the tau framework series
% Author: Sheng-Kai Huang
% Compile with: pdflatex paper2_gravity_tau.tex

\documentclass[prl,twocolumn,superscriptaddress,longbibliography,floatfix]{revtex4-2}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{bm}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{graphicx}

\newtheorem{theorem}{Theorem}

% Convenience macros (same as Paper 1 / Paper 1b)
\newcommand{\kB}{k_{\mathrm{B}}}
\newcommand{\Tr}{\mathrm{Tr}}
\newcommand{\id}{\mathrm{id}}
\newcommand{\Petz}{\mathcal{R}}
\newcommand{\chan}{\mathcal{N}}
\newcommand{\hilb}{\mathcal{H}}
\newcommand{\code}{\mathcal{C}}
\newcommand{\KL}{D}
\newcommand{\ket}[1]{|#1\rangle}
\newcommand{\bra}[1]{\langle#1|}
\newcommand{\braket}[2]{\langle#1|#2\rangle}
\newcommand{\op}[2]{|#1\rangle\langle#2|}
\newcommand{\abs}[1]{\left|#1\right|}
\newcommand{\norm}[1]{\left\|#1\right\|}
\newcommand{\tauasym}{\tau}

% Paper 2 specific macros
\newcommand{\ceff}{c_{\mathrm{eff}}}
\newcommand{\rs}{r_{\mathrm{s}}}
\newcommand{\Sgrav}{\Sigma_{\mathrm{grav}}}
\newcommand{\Fbound}{F_{\mathrm{bound}}}
\newcommand{\nref}{n_{\mathrm{grav}}}

\begin{document}

\title{Information Loss is Local:\\
The Gravitational Refractive Index as Petz Recovery Fidelity}

\author{Sheng-Kai Huang}
\email{akai@fawstudio.com}
\affiliation{Independent Researcher}

\date{\today}

\begin{abstract}
The temporal asymmetry parameter
$\tauasym = 1 - F \in [0,1]$, established in Paper~I
from the Petz recovery framework, quantifies the arrow
of time with the universal bound $F \geq e^{-\Sigma/2}$.
Defining the gravitational entropy production
$\Sgrav \equiv -\ln(-g_{00}) = \rs/r$, we obtain the
identification
$\ceff/c = e^{-\Sgrav/2} = \Fbound$:
the coordinate speed of light \emph{is} the Petz recovery
fidelity bound [\textsc{new synthesis}].
We then observe that $\tauasym(r) = 1 - \sqrt{-g_{00}(r)}$
depends only on the \emph{local} metric component $g_{00}$---no
knowledge of the global causal structure is needed.
This locality has a notable consequence:
the operationally relevant boundary where
$\tauasym = 1$ is the \emph{apparent horizon},
not the event horizon.
Using the Kodama vector for dynamical spacetimes, we show
that $\tauasym$ tracks the apparent horizon in Vaidya
collapse and the Hubble radius in FRW cosmology---consistent
with the quantum extremal surface prescription.
Since $\tauasym < 1$ everywhere for finite
$\Sgrav$, event horizons are excluded within this
framework: information that falls in must eventually
come out.
The exponential metric $g_{00} = -e^{-\rs/r}$,
which saturates the Petz bound, predicts
gravitational-wave echoes at
$\Delta t \approx 4\,\rs/c$---in principle testable
with existing LIGO--Virgo data, though not yet searched
for with appropriate templates.
\end{abstract}

\maketitle

%=======================================================================
\section{Introduction}
\label{sec:intro}
%=======================================================================

In Paper~I~\cite{Paper1}, we established that the Petz
recovery map~\cite{Petz1988}---the unique Bayesian
retrodiction functor~\cite{Parzygnat2023}---defines
a temporal asymmetry parameter
$\tauasym \equiv 1 - F(\rho,\,\Petz_{\sigma,\chan}\!\circ\!\chan(\rho))
\in [0,1]$, bounded by~\cite{Fawzi2015,JRSWW2018}
\begin{equation}
  \tauasym \;\leq\; 1 - e^{-\Sigma/2}\,,
  \label{eq:tau_bound_paper1}
\end{equation}
where $\Sigma$ is the entropy production of the
channel $\chan$.
$\tauasym = 0$ for closed (unitary) evolution;
$\tauasym > 0$ for open systems.

General relativity admits a well-known reformulation in which
gravity acts as a medium with position-dependent refractive
index~\cite{Dicke1957,Einstein1911}:
the coordinate speed of light is
$\ceff/c = e^{\Phi/c^2}$, where $\Phi < 0$ is the
Newtonian potential.
In this paper we show that
\begin{equation}
  \frac{\ceff}{c}
  \;=\; e^{\Phi/c^2}
  \;=\; e^{-\Sgrav/2}
  \;=\; \Fbound\,.
  \label{eq:preview}
\end{equation}
Light slowing down in a gravitational field \emph{is}
information loss; the Petz recovery fidelity \emph{is}
the refractive index.

This identification has a consequence that goes beyond
translation.
The quantity
$\tauasym(r) = 1 - \sqrt{-g_{00}(r)}$
is \emph{local}: to evaluate the degree of information
loss at a spacetime point, one needs only $g_{00}$
at that point---not the global causal structure, not
the future evolution of the spacetime.
In general relativity, event horizons are defined
teleologically: determining whether a point lies inside
an event horizon requires tracing null geodesics to
future null infinity, which presupposes knowledge of
the spacetime's entire future~\cite{Hawking1973}.
But $\tauasym$ makes no reference to the future.
The boundary where $\tauasym = 1$ is set by
\emph{local} data, and that boundary is the
\emph{apparent horizon}---a quasi-local,
causal surface~\cite{Hayward1994,AshtekarKrishnan2004}.

The logical chain of this paper is:
\begin{enumerate}
  \item $\ceff/c = \Fbound$
    (Sec.~\ref{sec:identification}: triple identification);
  \item $\tauasym$ is local
    (Sec.~\ref{sec:locality}: information loss does not
    require the future);
  \item $\tauasym < 1$ everywhere for finite $\Sgrav$
    (Sec.~\ref{sec:nohorizon}: no event horizon within
    this framework);
  \item echoes at $\Delta t \approx 4\,\rs/c$
    (Sec.~\ref{sec:predictions}: testable consequence).
\end{enumerate}

%=======================================================================
\section{The Triple Identification}
\label{sec:identification}
%=======================================================================

\subsection{The coordinate speed of light}

[\textsc{known}]  In the post-Newtonian approximation of GR,
the coordinate speed of light in a static gravitational field
is~\cite{Einstein1911,Dicke1957}
\begin{equation}
  \frac{\ceff}{c} = e^{\Phi/c^2}\,,
  \label{eq:ceff}
\end{equation}
where $\Phi < 0$ is the Newtonian gravitational potential.
This result, first anticipated by Einstein in
1911~\cite{Einstein1911} and placed on rigorous footing by
Dicke~\cite{Dicke1957}, is exact for the exponential
metric and agrees with the Schwarzschild metric through
$O(\rs/r)^2$.

\subsection{Gravitational entropy production}

[\textsc{new definition}]  We define the dimensionless
gravitational entropy production as
\begin{equation}
  \boxed{\;\Sgrav \;\equiv\; \frac{2|\Phi|}{c^2}
  \;=\; \frac{\rs}{r}\,,\;}
  \label{eq:Sgrav_def}
\end{equation}
where $\rs = 2GM/c^2$ is the Schwarzschild radius.
This is an \emph{informational} entropy production---the cost
of retrodiction from infinity to radius $r$---not a
thermodynamic entropy in the Clausius sense.
The Clausius entropy production for a photon in Tolman thermal
equilibrium vanishes identically~\cite{Tolman1930}, because
$T(\Phi)\sqrt{-g_{00}} = T_\infty$.
By contrast, $\Sgrav$ quantifies the informational cost of
climbing the gravitational potential, which persists even in
equilibrium.

Three independent arguments converge on
Eq.~\eqref{eq:Sgrav_def}:

\emph{(a) Modular flow}~[\textsc{known}].---In the
entanglement first-law regime of AQFT on Schwarzschild, the
fractional quantum relative entropy (QRE) loss between radius
$r$ and infinity is $(D_{\mathrm{in}} - D_{\mathrm{out}})/
D_{\mathrm{in}} = \rs/r$, exactly, via the Tolman blueshift
and energy redshift~\cite{DorauMuch2025}.

\emph{(b) Gravitational Landauer principle}~[\textsc{known}].---The
Tolman-modified cost of erasing one bit at radius $r$ exceeds
that at infinity by a factor $1/\sqrt{-g_{00}}$.
The dimensionless excess erasure cost yields
$\Sgrav = -\ln(-g_{00})$~\cite{Herrera2020}, which equals
$\rs/r$ for the exponential metric and
$-\ln(1 - \rs/r) \approx \rs/r + O((\rs/r)^2)$ for
Schwarzschild.

\emph{(c) Quantum channel model}~[\textsc{known}].---Modeling
gravitational redshift as a pure-loss bosonic channel with
intensity transmissivity $\eta = -g_{00}$, the QRE decrease
is $\Sigma = -\ln\eta = \rs/r$~\cite{AhmadiFuentes2014}.

All three routes share the common mathematical core
$\Sgrav = -\ln(-g_{00})$.

\subsection{The gravitational thermal attenuator channel}
\label{sec:grav_channel}

[\textsc{new result}]  The three routes above are not merely
analogies---they are facets of a single rigorous construction.

\begin{theorem}[Gravitational thermal attenuator]
\label{thm:grav_channel}
For a free scalar field mode of frequency $\omega$ propagating
radially from $r_1$ to $r_2 > r_1$ in a static metric
$ds^2 = -f(r)\,c^2\,dt^2 + f(r)^{-1}dr^2 + r^2 d\Omega^2$,
in the geometric optics limit
$(\omega^2 \gg V_{\mathrm{eff}})$:
\emph{(a)}~there exists a CPTP thermal attenuator
$\chan_{r_1 \to r_2}$ with transmissivity
$\eta = f(r_1)/f(r_2)$;
\emph{(b)}~the environment is thermal at the local Tolman
temperature $T_{\mathrm{loc}}(r_1) = T_H/\sqrt{f(r_1)}$;
\emph{(c)}~the entropy production satisfies
\begin{equation}
  \Sgrav(r) = -\ln f(r) + O(\omega^2/T^2)
  = -\ln(-g_{00})\,;
  \label{eq:Sgrav_channel}
\end{equation}
\emph{(d)}~the Petz recovery fidelity obeys
$F \geq e^{-\Sgrav/2} = \sqrt{f(r)}$;
\emph{(e)}~greybody corrections enter at
$O((\rs/r)^2)$~\cite{Wald1994}.
\end{theorem}

\emph{Proof sketch.}---The WKB mode equation gives amplitude
transmission $|t_\omega|^2 = f(r_1)/f(r_2)$: the energy ratio
(gravitational redshift) contributes $\sqrt{f(r_1)/f(r_2)}$
and the time-dilation ratio of proper times contributes an
identical factor~\cite{BirrellDavies1982}.
The Tolman relation identifies the environment as thermal,
yielding a Gaussian thermal attenuator
$\mathcal{E}_{\eta,N_B}$~\cite{IvanSabapathySimon2011} with
explicit Kraus operators on Fock space.
In the high-temperature limit,
$\Sigma = -\ln\eta + O(\omega^2/T^2)$~\cite{HolevoWerner2001},
establishing~(c).
Part~(d) follows from the JRSWW bound~\cite{JRSWW2018}.
The leading greybody correction is $|R_0|^2 \sim (\omega\rs)^2$,
giving $|\Sigma_{\mathrm{exact}} - \Sigma_0|
\leq \tfrac{1}{4}(\rs/r)^2$ (see Supplemental Material~\cite{supplement}).
\hfill$\square$

\medskip

Three corollaries follow immediately:
(i)~at the horizon ($f \to 0$), $\Sgrav \to \infty$,
$F \to 0$, $\tauasym \to 1$---complete information loss;
(ii)~in flat spacetime ($f = 1$), $\Sgrav = 0$, $F = 1$,
$\tauasym = 0$---no gravitational arrow of time;
(iii)~channels compose: $\Sgrav(r_1 \to r_3) =
\Sgrav(r_1 \to r_2) + \Sgrav(r_2 \to r_3)$.
The gravitational refractive index satisfies
$\nref = e^{\Sgrav/2} = 1/\Fbound$: \emph{refraction is
inverse recovery}~\cite{Dicke1957}.

This theorem places the $\tauasym$ framework on a
constructive footing:
$\tauasym$ is the recovery failure of a specific,
constructible CPTP map with explicit Kraus operators.

\subsection{The triple identification}

[\textsc{new synthesis}]  Substituting
Eq.~\eqref{eq:Sgrav_def} into the Petz fidelity bound
$F \geq e^{-\Sigma/2}$ from Paper~I:
\begin{equation}
  \Fbound = e^{-\Sgrav/2} = e^{\Phi/c^2}\,.
  \label{eq:Fbound}
\end{equation}
Comparing with Eq.~\eqref{eq:ceff}, we obtain the central
result:
\begin{equation}
  \boxed{\;\frac{\ceff}{c}
  \;=\; e^{-\Sgrav/2}
  \;=\; \Fbound\,.\;}
  \label{eq:triple}
\end{equation}
The effective speed of light \emph{is} the Petz recovery
fidelity bound.
Equivalently, the gravitational refractive index is the inverse
fidelity:
\begin{equation}
  \nref(r) = \frac{c}{\ceff} = e^{+\Sgrav/2} = \frac{1}{\Fbound}\,.
  \label{eq:refractive_index}
\end{equation}

Table~\ref{tab:numerical} provides numerical verification
across four orders of magnitude in gravitational compactness.

\begin{table}[t]
\caption{\label{tab:numerical}Gravitational entropy production
$\Sgrav = \rs/r$ and Petz recovery fidelity bound
$\Fbound = e^{-\Sgrav/2}$ for representative astrophysical
objects.  The last row indicates the limit of the weak-field
approximation.}
\begin{ruledtabular}
\begin{tabular}{lccc}
  \textbf{Object} & $\rs/r$ & $\Fbound$ & $\tauasym_{\max}$ \\
  \colrule
  Earth surface    & $1.4{\times}10^{-9}$ & $0.99999999930$
    & $7.0{\times}10^{-10}$ \\
  Sun surface      & $4.2{\times}10^{-6}$ & $0.9999979$
    & $2.1{\times}10^{-6}$ \\
  Neutron star     & $0.4$ & $0.819$
    & $0.181$ \\
  Schwarzschild $r\!=\!\rs$ & $1$ & $0.607$
    & $0.393$ \\
\end{tabular}
\end{ruledtabular}
\end{table}

\emph{Caveat.}---The identification~\eqref{eq:triple} is exact
for the exponential metric
$g_{00} = -e^{-\rs/r}$ and agrees with Schwarzschild through
$O(\rs/r)$.
At $O((\rs/r)^2)$, the two metrics differ:
the Schwarzschild expansion gives
$g_{00}^{\text{Schw}} = -1 + \rs/r - \rs^2/(2r^2) + \cdots$,
identical to the exponential expansion through this order.
The first discrepancy appears at $O((\rs/r)^3)$.
In the solar system ($\rs/r \sim 10^{-6}$), this cubic
correction is of order $10^{-18}$---far below current
measurement capability~\cite{Bertotti2003}.

%=======================================================================
\section{Information Loss is Local}
\label{sec:locality}
%=======================================================================

\subsection{$\tauasym$ depends only on local $g_{00}$}
\label{sec:tau_local}

[\textsc{new synthesis}]
The temporal asymmetry at a spacetime point $p$
at radius $r$ from a mass $M$ is
\begin{equation}
  \tauasym(r) = 1 - \sqrt{-g_{00}(r)}\,.
  \label{eq:tau_local}
\end{equation}
This is a \emph{local, operational} quantity:
an observer at $r$ determines $\tauasym$ by comparing
a local clock with a reference clock at infinity.
No information about the future evolution of the
spacetime, no knowledge of what happens at other radii,
and no global causal analysis is required.

This stands in sharp contrast with the event horizon.
In general relativity, the event horizon is defined as
\begin{equation}
  \mathcal{H}^{+} = \partial J^{-}(\mathcal{I}^{+})\,,
  \label{eq:EH_def}
\end{equation}
the boundary of the causal past of future null
infinity~\cite{Hawking1973}.
To determine whether a point lies inside the event
horizon, one must know the \emph{entire future} of the
spacetime---including whether the black hole eventually
evaporates, accretes more mass, or merges with another.
The event horizon is a teleological concept:
it is defined by where light rays \emph{will eventually}
reach, not by what is happening \emph{now}.

The $\tauasym$ framework replaces this global,
teleological criterion with a local, operational one.
Information loss is not a verdict delivered by the
future---it is a condition determined by the present.

\subsection{The apparent horizon as $\tauasym = 1$ surface}
\label{sec:apparent_horizon}

[\textsc{new synthesis}]
In general relativity, the \emph{apparent horizon}
is defined as the outermost marginally outer trapped
surface (MOTS): the surface where the expansion
of outgoing null geodesics vanishes,
$\theta_{+} = 0$~\cite{Hayward1994}.
Unlike the event horizon, the apparent horizon
is quasi-local: it depends only on the metric and its
first derivatives on a single time-slice.

The $\tauasym$ framework naturally selects this surface.
For the Schwarzschild metric, $\tauasym = 1$ when
$g_{00} = 0$, which occurs at $r = \rs$---both the
event horizon and the apparent horizon coincide.
But in dynamical spacetimes they differ, and $\tauasym$
tracks the apparent horizon.
This has a precise mathematical content:
\begin{equation}
  \tauasym = 1 \;\;\Longleftrightarrow\;\;
  g_{00} = 0 \;\;\Longleftrightarrow\;\;
  \theta_{+} = 0 \;\text{(MOTS)}\,.
  \label{eq:tau_MOTS}
\end{equation}
The equivalence on the right holds for spherically
symmetric spacetimes where
$\theta_{+} = (2/r)(1 - 2M(v)/r)^{1/2}$:
$\theta_{+} = 0$ precisely when $r = 2M(v)$,
i.e., when $g_{00}^{\text{eff}} = 0$.

\emph{Connection to QES.}---[\textsc{new synthesis}]
The quantum extremal surface (QES) prescription of
Engelhardt and Wall~\cite{Engelhardt2015}
and its applications by
Penington~\cite{Penington2020} and the island
formula~\cite{AEMM2021} identify the QES---not the
classical event horizon---as the surface that controls
the entanglement entropy of Hawking radiation.
The QES extremizes the \emph{generalized} entropy
$S_{\text{gen}} = A/(4G\hbar) + S_{\text{bulk}}$
on a time-slice.
The $\tauasym$ framework provides a complementary
characterization: the QES is the surface where the
\emph{operational} information loss saturates
($\tauasym \to 1$), while the generalized entropy
formulation controls the \emph{counting} of degrees
of freedom.
Both agree that the operationally relevant boundary
is quasi-local and causal---defined by data on a single
slice, not by the global causal structure.
In the semiclassical limit, the QES reduces to the
apparent horizon~\cite{Engelhardt2015}, and $\tauasym = 1$
selects the same surface---the two prescriptions coincide.

\subsection{Dynamic extension: the Kodama vector}
\label{sec:kodama}

[\textsc{new synthesis}]
The static formulation of $\tauasym$ requires a
timelike Killing vector $\xi^{a}$.
The natural dynamical extension uses the Kodama
vector~\cite{Kodama1980}---the unique divergence-free
timelike vector field in spherically symmetric
spacetimes:
\begin{equation}
  K^{a} = \epsilon^{ab}\nabla_{b} R\,,
  \label{eq:kodama_def}
\end{equation}
where $R$ is the areal radius and $\epsilon^{ab}$ is the
Levi-Civita tensor on the $(t,r)$ submanifold.
In the static limit, $K^{a}$ reduces to the Killing vector
$\xi^{a}$, and $|K| = \sqrt{-g_{00}}$.
The Kodama recovery fidelity is
\begin{equation}
  \Fbound^{K}(A\!\to\! B) = \frac{|K|_{A}}{|K|_{B}}\,,
  \label{eq:FK}
\end{equation}
recovering all results of Sec.~\ref{sec:identification}
in the static limit.

\emph{Vaidya collapse.}---[\textsc{new synthesis}]
The ingoing Vaidya metric describes a spherically symmetric
spacetime with a time-dependent mass function $M(v)$:
\begin{equation}
  ds^{2} = -\!\left(1 - \frac{2M(v)}{r}\right)dv^{2}
  + 2\,dv\,dr + r^{2}\,d\Omega^{2}\,,
  \label{eq:vaidya_metric}
\end{equation}
where $v$ is the advanced (Eddington--Finkelstein) time
and $dM/dv > 0$ for accretion~\cite{Hiscock1981}.
The Kodama vector norm
is~\cite{Hayward1994,AbreuVisser2010}
$|K| = \sqrt{1 - 2M(v)/r}$, giving
\begin{equation}
  \tauasym_{K}(r,v) = 1 - \sqrt{1 - \frac{2M(v)}{r}}\,.
  \label{eq:tau_vaidya}
\end{equation}
This reaches $\tauasym = 1$ when $r = 2M(v)$---precisely
the \emph{apparent} horizon, the outermost MOTS on
each $v = \mathrm{const}$ slice.
During gravitational collapse, $M(v)$ increases and the
$\tauasym = 1$ surface expands outward, tracking the
growing apparent horizon.
During Hawking evaporation, $M(v)$ decreases and the
$\tauasym = 1$ surface shrinks---information recovery
improves, consistent with the Page curve.

\emph{Explicit demonstration: AH $\neq$ EH.}---The
event horizon is the outermost outgoing null geodesic that
fails to escape to infinity.
For radial null rays in Vaidya,
$dr/dv = \frac{1}{2}(1 - 2M(v)/r)$, integrated backward
from $r_{\text{EH}} = 2M_0$ at $v \to \infty$.
For the self-similar profile $M(v) = \lambda v$, the null
geodesic equation admits the exact solution
$r_{\text{EH}}(v) = C_{-}\,v$, where
$C_{-} = (1 - \sqrt{1 - 16\lambda})/4$, while
$r_{\text{AH}}(v) = 2\lambda v$.
Since $C_{-} > 2\lambda$ for $\lambda > 0$
(see Supplemental Material~\cite{supplement} for proof),
$r_{\text{EH}} > r_{\text{AH}}$ during accretion:
the event horizon lies \emph{outside} the apparent horizon.

Table~\ref{tab:AH_EH} quantifies the discrepancy.
At the event horizon, $\tauasym_{K} < 1$---the
$\tauasym$ framework says the EH is \emph{not} a surface
of complete information loss.
At the apparent horizon, $\tauasym_{K} = 1$ exactly,
for all $v$.
During evaporation ($dM/dv < 0$), the situation reverses:
$r_{\text{EH}} < r_{\text{AH}}$, and the Kodama vector
becomes spacelike at the EH.
In both cases, $\tauasym_{K} = 1$ at the AH---never
at the EH.

\begin{table}[t]
\caption{\label{tab:AH_EH}%
$\tauasym_{K}$ at the event horizon versus the
apparent horizon for the self-similar Vaidya
spacetime $M(v) = \lambda v$.
$\tauasym_{K}(\text{AH}) = 1$ exactly for all
$\lambda$; $\tauasym_{K}(\text{EH}) < 1$.
See Supplemental Material for derivations.}
\begin{ruledtabular}
\begin{tabular}{cccc}
  $\lambda$ & $r_{\text{EH}}/r_{\text{AH}}$
    & $\tauasym_{K}(r_{\text{EH}})$
    & $1 - \tauasym_{K}(r_{\text{EH}})$ \\
  \colrule
  $0.001$ & $1.004$ & $0.937$ & $6.3\%$ \\
  $0.01$  & $1.044$ & $0.796$ & $20.4\%$ \\
  $0.03$  & $1.162$ & $0.627$ & $37.3\%$ \\
  $0.05$  & $1.382$ & $0.474$ & $52.6\%$ \\
  $1/16$  & $2.000$ & $0.293$ & $70.7\%$ \\
\end{tabular}
\end{ruledtabular}
\end{table}

For a step-function collapse ($M(v) = M_0\,\theta(v)$),
the event horizon originates at $r = 0$ when
$v = -4M_0$---in the \emph{flat} region, before the
shell arrives.
The EH ``anticipates'' the collapse; the AH appears
only at $v = 0$.
This illustrates the distinction clearly:
$\tauasym_{K}$ is zero in the flat region (no information
loss), even though the event horizon is already present.

\emph{FRW cosmology.}---In a spatially flat FRW universe
with Hubble parameter $H$,
\begin{equation}
  \tauasym_{K}(R) = 1 - \sqrt{1 - H^{2}R^{2}}\,,
  \label{eq:tau_FRW}
\end{equation}
reaching $\tauasym = 1$ at $R = 1/H$, the Hubble
radius---which is the apparent horizon of FRW, not an
event horizon~\cite{AbreuVisser2010}.

\subsection{Verification proposal: SYK model}
\label{sec:SYK}

[\textsc{speculative}]
The SYK model~\cite{SachdevYe1993,MaldacenaStanford2016}
is a solvable quantum-mechanical model with a holographic
dual to nearly-AdS$_{2}$ (JT) gravity, exhibiting both
scrambling and black hole formation.
It provides a setting where both sides of the $\tauasym$
equation---the QI recovery fidelity $F(t)$ and the
gravitational entropy production $\Sgrav(t)$---can be
computed independently.

We propose the following test.
In the SYK model with time-dependent coupling (modeling
dynamical black hole formation and evaporation):
(i)~compute the Petz recovery fidelity $F(t)$
for an early-time operator from late-time data;
(ii)~compute $\Sgrav(t) = -\ln(-g_{00}^{\text{bulk}}(t))$
from the dual JT gravity;
(iii)~compare whether $F(t)$ tracks the apparent
horizon or the event horizon in the bulk.
The $\tauasym$ framework predicts that $F(t)$ tracks the
apparent horizon.
This calculation is feasible with exact diagonalization
at moderate $N$~\cite{MaldacenaStanford2016}; the
prediction awaits numerical verification.

%=======================================================================
\section{No Event Horizon Within the $\tauasym$ Framework}
\label{sec:nohorizon}
%=======================================================================

\subsection{Model-independent no-horizon theorem}

[\textsc{new synthesis}]
The locality of $\tauasym$ established in
Sec.~\ref{sec:locality} has a direct physical consequence.
If gravitational redshift is a quantum channel with
\emph{finite} entropy production $\Sigma < \infty$, then
the Petz bound gives
$F \geq e^{-\Sigma/2} > 0$, so $\tauasym < 1$.
An event horizon requires $F = 0$ (complete information loss),
which demands $\Sigma \to \infty$---impossible for a
finite-mass object at finite distance.
Under this premise, the Petz bound~\cite{JRSWW2018} excludes
event horizons; neither the specific
form of $\Sgrav$ nor the saturation hypothesis is needed.

In the language of Sec.~\ref{sec:locality}: because
$\tauasym$ is local and finite, information loss is
never complete.
Within this framework, what falls in must eventually come out.

\subsection{The exponential metric}

[\textsc{new synthesis}]
If the Petz bound is taken as an identity---$\sqrt{-g_{00}}
= e^{-\Sgrav/2}$---the metric component is uniquely
determined.
This gives the exponential
metric~\cite{Papapetrou1954,Yilmaz1958}:
\begin{equation}
  ds^2 = -e^{-\rs/r}\, c^2 dt^2
       + e^{+\rs/r}\!\left(dr^2 + r^2 d\Omega^2\right),
  \label{eq:exp_metric}
\end{equation}
in isotropic coordinates.
Since $e^{-\rs/r} > 0$ for all $r > 0$,
$\tauasym < 1$ everywhere.

\emph{Exact solution.}---The exponential metric is not a vacuum
solution of the Einstein equations ($R_{\mu\nu} \neq 0$).
It is an exact solution of Einstein gravity coupled to a phantom
scalar field $\phi = M/r$~\cite{ErnazarovIvashchuk2026}, and
also arises in logarithmic $f(R)$
gravity~\cite{NathSarma2025} and
scalar-Einstein-Gauss-Bonnet theory~\cite{Ernazarov2025}.
Makukov and Mychelkin~\cite{Makukov2020} showed that three
independent scalar-field families within GR all reduce to
the exponential metric when the scalar charge equals the
gravitational mass.
Boonserm et~al.~\cite{Boonserm2018} classified the geometry as
a traversable wormhole.

\emph{The spatial metric is assumed, not derived.}---The
information-theoretic framework constrains only the temporal
component $g_{00}$ through $\sqrt{-g_{00}} = e^{-\Sgrav/2}$.
The spatial component $g_{rr} = e^{+\rs/r}$ in
Eq.~\eqref{eq:exp_metric} follows from the isometry condition
$g_{00}\cdot g_{rr} = -1$ (isotropic coordinates), which is
equivalent to Dicke's scalar refractive
index~\cite{Dicke1957}: $n^2 = 1/(-g_{00})$ with $g_{rr} = n^2$.
This is a physically motivated assumption---not a derivation from
quantum information theory.
The fundamental reason is that quantum channels model
\emph{temporal} processes (evolution, decoherence, loss),
and the Petz recovery framework naturally constrains the
temporal metric component; no analogous information-degradation
process exists for the spatial direction.
Consequently, we classify our predictions as
\emph{robust} (depending only on $g_{00}$: no-horizon theorem,
refractive index, gravitational decoherence rate) or
\emph{conditional} (depending on the full metric:
shadow size, ISCO radius, QNM frequencies, echo delay).
The Jacobson~\cite{Jacobson1995} and
Dorau--Much~\cite{DorauMuch2025} programs \emph{do} derive
all metric components from information-theoretic premises,
but their vacuum solution is Schwarzschild, not exponential.
More recently, Thattarampilly et~al.~\cite{Thattarampilly2026}
solved the full Gravity-from-Entropy (GfE) field equations of
Bianconi~\cite{Bianconi2025} for the static spherically
symmetric vacuum, finding Schwarzschild with $O(r^{-4})$
corrections---again not the exponential metric.
Every known entanglement-based gravity program thus produces
Schwarzschild (plus perturbative corrections) as its vacuum
solution.
This gap remains an open question: the exponential metric is
best understood as a \emph{Petz saturation ansatz}---the
geometry defined by promoting $F = e^{-\Sgrav/2}$ to an
identity---rather than the output of any established
entropy-to-gravity derivation.

\emph{Comparison with Schwarzschild.}---A persistent
misconception holds that the exponential metric ``predicts
half the observed light bending.''
This conflates the Yilmaz \emph{field equations}---an
alternative theory of gravity shown to be inconsistent by
Misner~\cite{Misner1999}---with the exponential \emph{metric}
as a solution of standard GR coupled to a phantom scalar.
Particles and photons follow standard geodesics
of~\eqref{eq:exp_metric}; only the geometry differs from
Schwarzschild.

The PPN expansion in isotropic coordinates gives
\begin{align}
  g_{00}^{\text{exp}} &= -e^{-\rs/r}
  = -(1 - 2U + 2U^2 - \cdots) \,,\\
  g_{ij}^{\text{exp}} &= e^{+\rs/r}\,\delta_{ij}
  = (1 + 2U + 2U^2 + \cdots)\,\delta_{ij}\,,
\end{align}
where $U = \rs/(2r)$ is the Newtonian potential.
Matching with the standard PPN form
$g_{00} = -(1-2U+2\beta U^2+\cdots)$,
$g_{ij} = (1+2\gamma U+\cdots)\delta_{ij}$
yields $\gamma = \beta = 1$---\emph{identical to GR}.
Mychelkin et~al.~\cite{Mychelkin2024} verified explicitly
that the perihelion precession at first post-Newtonian order
matches Schwarzschild:
$\Delta\phi_{\text{exp}} = \Delta\phi_{\text{Schw}}
= 6\pi M/p$.
The first difference appears at 2PN, where the precession
differs by $\sim 2.3\times 10^{-7}$~arcsec/century for
Mercury---four orders of magnitude below current precision.
All solar-system tests are passed: light bending, Shapiro
delay, perihelion precession, and the
Cassini bound $|\gamma - 1| < 2.3\times 10^{-5}$~%
\cite{Bertotti2003}.
\emph{The exponential metric is observationally degenerate
with Schwarzschild for every current and foreseeable
weak-field test.}

\emph{Caveat on saturation.}---The JRSWW
bound~\cite{JRSWW2018} is structurally strict for
$\Sigma > 0$: the proof via the Golden--Thompson inequality
is an equality only when the relevant operators
commute---the $\Sigma = 0$ case~\cite{Petz1988}.
No known quantum channel saturates the bound for
$\Sigma > 0$.
The exponential metric should be understood as the geometry
defined by the \emph{identity}
$\sqrt{-g_{00}} = e^{-\Sgrav/2}$; whether nature saturates
this bound or admits a gap is empirical.

%=======================================================================
\section{Strong-Field Predictions}
\label{sec:predictions}
%=======================================================================

[\textsc{conditional}]
In the strong-field regime, the exponential and Schwarzschild
metrics diverge, yielding three testable predictions.
All depend on the spatial metric assumption
$g_{rr} = e^{+\rs/r}$.

\emph{(1) Shadow size.}---The exponential photon sphere
at $r_{\text{ph}} = \rs$ (isotropic coordinates)
yields a critical impact parameter
$b_{\text{crit}} = 2e\,m \approx 5.44\,m$
versus Schwarzschild's $3\sqrt{3}\,m \approx 5.20\,m$
($m = GM/c^2$): a $4.6\%$ larger shadow.
The EHT measurement of $42 \pm 3\;\mu\text{as}$ for
M87*~\cite{EHT2019} is consistent with both metrics.
This prediction awaits next-generation VLBI observations
with sufficient angular resolution to distinguish the two.

\emph{(2) Quasinormal modes.}---In the eikonal limit,
$f_{\text{QNM}}^{\text{exp}}/
f_{\text{QNM}}^{\text{Schw}}
= 3\sqrt{3}/(2e) \approx 0.956$,
a $\sim 4.4\%$ downward shift.
WKB calculations find a $-7.8\%$ shift for a generalized
exponential wormhole~\cite{NathSarma2025}.
This prediction has not yet been tested; next-generation
gravitational-wave detectors
(Einstein Telescope, Cosmic Explorer) are expected to reach
$\sim 1\%$ precision in QNM frequencies.

\emph{(3) Gravitational-wave echoes}~[\textsc{most
discriminating prediction}].---A concrete and
falsifiable signature of the $\tauasym$ framework.
The traversable wormhole structure of the exponential
metric produces echoes---repeated, damped pulses following
the initial ringdown.
The echo delay is set by the round-trip light travel time
in the tortoise coordinate $r^* = \int e^{\rs/r}\,dr$.
For the exponential metric, $r^*$ remains bounded
(no logarithmic divergence as in Schwarzschild), and
the effective potential cavity extends from the photon
sphere at $r_{\mathrm{ph}} = \rs$ (isotropic) to the
wormhole throat at $r_{\mathrm{throat}} = \rs/2$.
The echo delay is
\begin{equation}
  \Delta t_{\text{echo}}
  = \frac{2}{c}\!\int_{r_{\mathrm{throat}}}^{r_{\mathrm{ph}}}
  \!\!e^{\rs/r}\,dr
  \;\approx\; \frac{4\,\rs}{c}
  \ln\!\left(\frac{\rs}{r_{\mathrm{throat}}}\right),
  \label{eq:echo}
\end{equation}
where the logarithmic approximation holds because
$e^{\rs/r}$ is sharply peaked near the throat.
Numerically, the exact integral gives
$\Delta t_{\text{echo}} \approx 4.2\,\rs/c$.
Table~\ref{tab:echo} gives explicit estimates for
astrophysical sources across the LIGO and LISA bands.

\begin{table}[t]
\caption{\label{tab:echo}%
Echo delay $\Delta t_{\text{echo}} \approx 4.2\,\rs/c$
for representative sources compared with Planck-wall
models~\cite{Cardoso2016} where
$\Delta t_{\text{Planck}} \approx \rs\ln(\rs/\ell_P)/c$.
The exponential-metric echoes are $\sim 2$ orders of
magnitude shorter.}
\begin{ruledtabular}
\begin{tabular}{lccc}
  \textbf{Source} & $M$ & $\Delta t_{\text{echo}}$
    & $\Delta t_{\text{Planck}}$ \\
  \colrule
  Stellar BH (GW150914) & $60\,M_\odot$
    & $2.5$~ms & $\sim 230$~ms \\
  Intermediate-mass BH & $10^{3}\,M_\odot$
    & $40$~ms & $\sim 4$~s \\
  Sgr A* & $4{\times}10^{6}\,M_\odot$
    & $165$~s & $\sim 1.9{\times}10^{4}$~s \\
  M87* & $6.5{\times}10^{9}\,M_\odot$
    & $74$~hr & $\sim 350$~d \\
\end{tabular}
\end{ruledtabular}
\end{table}

The $2.5$~ms echo for stellar-mass mergers falls in
the LIGO--Virgo--KAGRA sensitivity band
($20$--$2000$~Hz; the echo frequency
$\sim 1/\Delta t \sim 400$~Hz is near peak sensitivity).
For supermassive black holes, the echo timescale of
minutes to hours is within the LISA band
($10^{-4}$--$10^{-1}$~Hz).
This is $\sim 90\times$ shorter than Planck-wall
models~\cite{Cardoso2016}; current echo searches use
$\sim 200$~ms templates appropriate for Planck-wall
scenarios and would \emph{miss} the exponential-metric
signal.
Dedicated short-delay searches in existing
LIGO--Virgo O3/O4 data would be valuable.

The echo prediction follows directly from the no-horizon
theorem: since $\tauasym < 1$ everywhere, the
gravitational potential forms a well, not a one-way
membrane.
Gravitational waves enter, scatter off the wormhole
throat, and return---carrying the signature of
incomplete information loss.
This echo prediction has not yet been tested with
appropriate short-delay templates.
A null detection with templates matched to
Eq.~\eqref{eq:echo} would falsify the saturation
hypothesis ($\sqrt{-g_{00}} = e^{-\Sgrav/2}$), while
a positive detection would provide direct evidence
for the exponential metric.

\emph{Summary of falsifiable predictions.}---Table~\ref{tab:predictions_summary}
collects the three concrete predictions with numerical values.
All three await experimental verification.

\begin{table}[t]
\caption{\label{tab:predictions_summary}%
Falsifiable predictions of the exponential metric
(conditional on $g_{rr} = e^{+\rs/r}$).
None has been tested to date.}
\begin{ruledtabular}
\begin{tabular}{lcc}
  \textbf{Observable} & \textbf{Prediction}
    & \textbf{Schwarzschild} \\
  \colrule
  Shadow $b_{\text{crit}}/m$ & $2e \approx 5.44$
    & $3\sqrt{3} \approx 5.20$ \\
  QNM shift & $-4.4\%$ & (reference) \\
  Echo delay (stellar) & $\approx 2.5$~ms
    & none (horizon) \\
\end{tabular}
\end{ruledtabular}
\end{table}

\emph{Honest assessment.}---Since both metrics share
$\gamma = \beta = 1$, weak-field data cannot
distinguish them; the first differences are
inherently \emph{strong-field} effects.
If nature selects $F > e^{-\Sgrav/2}$, the actual
metric lies between the exponential and Schwarzschild
predictions.
We emphasize a two-layer structure:
the \emph{robust} predictions (no-horizon theorem,
$\nref = 1/\Fbound$, decoherence bound) depend only on
$g_{00}$ and hold regardless of the spatial metric;
the \emph{conditional} predictions (shadow, QNM,
echoes) depend on the isometry assumption
$g_{rr} = e^{+\rs/r}$.
The most characteristic signature---short-delay
echoes~\eqref{eq:echo}---has not yet been searched
for with appropriate templates in LIGO--Virgo data.

%=======================================================================
\section{Supporting Connections}
\label{sec:connections}
%=======================================================================

The $\tauasym$ framework connects to several established
results in gravitational physics.
We classify each connection in Table~\ref{tab:classification}.

\emph{Gravitational decoherence.}---[\textsc{new synthesis}]
Pikovski et~al.~\cite{Pikovski2015} showed that gravitational
time dilation universally dephases composite quantum systems.
In the $\tauasym$ framework~\cite{Paper1b},
\begin{equation}
  \tauasym_{\mathrm{grav}}(t)
  = 1 - \exp\!\big(-N\Gamma t^2\big)\,,
  \label{eq:tau_pikovski}
\end{equation}
where $\Gamma = (\Delta E)^2(\Delta\Phi)^2/(2\hbar^2 c^4)$.
The dephasing rate is controlled by
$\Delta\Phi/c^2 \sim \Sgrav/2$: the gravitational entropy
production governs both the static (refractive index) and
dynamic (decoherence) aspects of information loss.

\emph{Unruh effect.}---[\textsc{new synthesis}]
A uniformly accelerating observer at acceleration $a$
perceives a thermal bath at $T_U = \hbar a/(2\pi c\kB)$,
inducing entropy production
$\Sigma_{\text{Unruh}} = 2\pi a\Delta x/(\hbar c^2)$
on any carried system~\cite{Unruh1976,LandiPaternostro2021}.
By the equivalence principle, this connects to $\Sgrav$.

\emph{Bekenstein bound.}---[\textsc{new synthesis}]
The bound $S \leq 2\pi ER/(\hbar c)$~\cite{Bekenstein1981}
translates to $\tauasym \leq 1 - \exp(-\pi ER/(\hbar c))$,
providing a universal upper limit on temporal asymmetry
for spatially bounded systems.

\emph{Complementarity.}---[\textsc{new synthesis}]
The infalling observer experiences
$\tauasym_{\text{in}} \approx 0$ (locally flat), while the
external observer describes
$\tauasym_{\text{ext}} \to 1$~\cite{Susskind1993}.
Different observer foliations assign different $\tauasym$
values---complementarity is observer-dependent $\tauasym$.

\emph{Page curve.}---[\textsc{new synthesis}]
The Page curve~\cite{Page1993} translates to
$\tauasym(t) \leq 1 - e^{-S_{\text{Page}}(t)/2}$:
$\tauasym$ grows before the Page time (information loss
worsens) and shrinks after it (recovery
improves)~\cite{Penington2020,AEMM2021}.

\emph{Hayden--Preskill.}---[\textsc{known}]
Cotler et~al.~\cite{Cotler2019} proved that the HP
decoder~\cite{HaydenPreskill2007} \emph{is} the Petz
recovery map, with fidelity satisfying the same
bound~\eqref{eq:tau_bound_paper1}.

\emph{Entanglement wedge.}---[\textsc{known}]
In AdS/CFT, bulk reconstruction via the Petz
map~\cite{ADH2015,CPS2020} satisfies
$F^2 \geq e^{-\Delta S_{\text{gen}}}$~\cite{Cotler2019}---the
same universal recovery inequality.

\begin{table}[t]
\caption{\label{tab:classification}Classification of results.}
\begin{ruledtabular}
\begin{tabular}{lc}
  \textbf{Connection} & \textbf{Status} \\
  \colrule
  $\ceff/c = \Fbound$ (refractive index)
    & \textsc{new synthesis} \\
  $\Sgrav = -\ln(-g_{00})$ (entropy production)
    & \textsc{new definition} \\
  $\tauasym$ is local $\to$ apparent horizon
    & \textsc{new synthesis} \\
  Kodama extension to dynamic spacetimes
    & \textsc{new synthesis} \\
  No-horizon theorem
    & \textsc{new synthesis} \\
  Observer-dependent $\tauasym$ (complementarity)
    & \textsc{new synthesis} \\
  Page curve as $\tauasym(t)$
    & \textsc{new synthesis} \\
  Gravitational decoherence as Petz recovery
    & \textsc{new synthesis} \\
  HP decoder = Petz map~\cite{Cotler2019}
    & \textsc{known} \\
  Entanglement wedge via Petz~\cite{CPS2020}
    & \textsc{known} \\
  $F^2 \geq e^{-\Delta S_{\mathrm{gen}}}$~\cite{Penington2020}
    & \textsc{known} \\
  SYK verification of $\tauasym$ tracking AH
    & \textsc{speculative} \\
\end{tabular}
\end{ruledtabular}
\end{table}

%=======================================================================
\section{What $\tauasym$ Can and Cannot Explain}
\label{sec:scope}
%=======================================================================

\emph{Can:}
\begin{itemize}
\item Identify information loss as a local, operational
  quantity that does not require knowledge of the future.
\item Identify the apparent horizon---not the event
  horizon---as the operationally relevant boundary,
  consistent with the QES prescription.
\item Unify quantum error correction, thermodynamic
  irreversibility, black hole information, and gravitational
  decoherence under a single parameter $\tauasym$.
\item Provide universal bounds: any gravitational channel
  satisfies $F \geq e^{-\Sgrav/2}$, constraining recovery
  independently of the specific protocol.
\item Derive the no-horizon theorem: $\tauasym < 1$ for any
  finite-mass object at finite distance
  (model-independent~\cite{JRSWW2018}).
\item Extend to dynamical spacetimes via the Kodama vector,
  tracking apparent horizons in Vaidya and FRW.
\end{itemize}

\emph{Cannot:}
\begin{itemize}
\item Explain dark matter: at galactic scales,
  $\Sgrav \sim 10^{-6}$, far too small to produce observable
  effects beyond Newtonian gravity.
\item Explain dark energy or the cosmological constant.
\item Derive $g_{ab}$ from $\tauasym$ alone: the relation is
  one-directional ($g \to \Sigma \to \tauasym$); $\tauasym$
  constrains only $g_{00}$.
  The spatial component $g_{rr}$ follows from the Dicke
  isometry condition; programs that derive all components
  from entanglement equilibrium
  (Jacobson, Dorau--Much, Bianconi GfE~\cite{Bianconi2025})
  yield Schwarzschild (plus perturbative
  corrections~\cite{Thattarampilly2026}) as the
  \emph{vacuum} solution.
  The exponential metric requires a phantom scalar source,
  whose information-theoretic origin remains open.
\item Provide the full non-perturbative Bogoliubov calculation
  beyond geometric optics: the channel of
  Theorem~\ref{thm:grav_channel} is derived in the WKB limit,
  with greybody corrections bounded at $O((\rs/r)^2)$.
\item Resolve the firewall paradox~\cite{AMPS2013}: while
  $\tauasym < 1$ removes the event horizon premise, the
  framework does not address the interior structure of compact
  objects.
\end{itemize}

%=======================================================================
\section{Discussion and Outlook}
\label{sec:discussion}
%=======================================================================

The central result of this paper is that gravitational
information loss is \emph{local}.
The temporal asymmetry $\tauasym = 1 - \sqrt{-g_{00}}$
depends only on the metric component at a single point.
The operationally relevant boundary for information loss is
therefore the apparent horizon---a quasi-local, causal
surface~\cite{AshtekarKrishnan2004}---not the event
horizon.
This is consistent with the quantum extremal surface
prescription~\cite{Penington2020,Engelhardt2015} and the
crossed-product algebra approach~\cite{Witten2022,%
Chandrasekaran2023}, which makes gravitational entropy
well-defined and observer-dependent~\cite{KudlerFlam2025}.

The identification $\ceff/c = \Fbound$
(Eq.~\eqref{eq:triple}) is complementary to the Jacobson
program~\cite{Jacobson1995}: while Jacobson derives
\emph{field equations} from thermodynamic entropy, we derive
the \emph{metric} directly from information recoverability.
The Dorau--Much result~\cite{DorauMuch2025}---deriving the
semiclassical Einstein equations from quantum relative
entropy---suggests that the two programs share a common
informational foundation.
Cirafici~\cite{Cirafici2024} has established fluctuation
theorems in gravitational algebras, confirming that the
Crooks--Petz connection extends to de~Sitter spacetime.

We emphasize the logical structure of this paper.
The results of Secs.~\ref{sec:identification}--\ref{sec:nohorizon}
---the triple identification, locality of $\tauasym$,
and the no-horizon theorem---hold for \emph{any} static metric
satisfying $\Sgrav = -\ln(-g_{00})$ with $\Sgrav < \infty$;
they are independent of whether the metric is Schwarzschild,
exponential, or the Schwarzschild-plus-corrections solution of
Gravity-from-Entropy~\cite{Thattarampilly2026,Bianconi2025}.
The exponential metric enters only in Sec.~\ref{sec:predictions},
where it serves as a maximal-recovery ansatz that saturates
$F = e^{-\Sgrav/2}$.
All currently known entanglement-based gravity programs
(Jacobson, Dorau--Much, Bianconi GfE) produce Schwarzschild
as the leading-order vacuum solution, with perturbative
corrections~\cite{Thattarampilly2026}.
Whether nature saturates the Petz bound
or admits a gap of order $O((\rs/r)^2)$ is empirical---and
the echo prediction of Sec.~\ref{sec:predictions} is precisely
the observable that distinguishes the two possibilities.

Several directions remain.
\emph{(i)~Beyond geometric optics}: the gravitational thermal
attenuator (Theorem~\ref{thm:grav_channel}) is established in
the WKB limit; computing the exact greybody spectrum for the
exponential metric would determine whether the channel
\emph{saturates} the Petz bound or admits a gap of order
$O((\rs/r)^2)$.
\emph{(ii)~Rotation}: the Kerr analog of the exponential
metric~\cite{Graber2017,MakukovMychelkin2023} lacks both
horizons and ergospheres; extending the framework beyond
spherical symmetry requires the axisymmetric
Kodama-type vector of Dorau and
Verch~\cite{DorauVerch2024}.
\emph{(iii)~Entanglement-assisted recovery}:
[\textsc{new synthesis}]  When the gravitational channel
$\chan_{\mathrm{grav}}$ acts on a system $L$ that is entangled
with a reference $R$, the total QRE decrease decomposes as
$\Delta D_{\mathrm{total}} = \Sgrav + I(L;R) - I(L';R) \geq 0$
(by the data-processing inequality), where $L' = \chan(L)$.
The single-party bound $F \geq e^{-\Sgrav/2}$ is derived
without access to $R$; when the recovery map has access to the
quantum side information in $R$, the assisted fidelity satisfies
$F_{\mathrm{assisted}} \geq
e^{-(\Sgrav + I(L;R) - I(L';R))/2}$.
If the channel approximately preserves entanglement
($I(L';R) \approx I(L;R)$), then
$F_{\mathrm{assisted}} \gg e^{-\Sgrav/2}$:
the recovery far exceeds the single-party geometric bound.
This is precisely the mechanism behind the Gao--Jafferis--Wall
traversable wormhole~\cite{GaoJafferisWall2017} and the
Hayden--Preskill protocol~\cite{HaydenPreskill2007}---shared
entanglement provides a quantum side channel that enhances
recovery without making the total entropy production negative.
The assisted vs.\ unassisted gap
$\Delta F \equiv F_{\mathrm{assisted}} - e^{-\Sgrav/2}$
is a continuous measure that unifies the Page curve transition
(growing $I(L';R)$) with wormhole traversability
(high $I(L';R)/I(L;R)$).
This prediction is testable via SYK numerics and quantum
circuit experiments~\cite{Jafferis2022}.
\emph{(iv)~Non-Markovian gravity}: if the gravitational
channel exhibits memory effects, $\Sgrav$ may not be additive
and the Petz bound requires
modification~\cite{LandiPaternostro2021}.
This could serve as a signature of quantum gravity:
classical gravity is necessarily Markovian, while quantum
gravity may permit non-Markovian information backflow.

In Paper~1b~\cite{Paper1b}, we applied the $\tauasym$ framework
to wavefunction collapse.
A future paper (Paper~3) will investigate the weak-field limit
at galactic scales.

\begin{acknowledgments}
The author thanks the quantum information, quantum gravity,
and quantum foundations communities for the theoretical
foundations on which this work builds.
\end{acknowledgments}

%=======================================================================
% BIBLIOGRAPHY
%=======================================================================

\begin{thebibliography}{55}

% === Paper series ===

\bibitem{Paper1}
S.-K. Huang,
``Universal Recovery: The Petz Map as Retrodiction Functor,
Error Corrector, and Entropy Bound,''
companion paper (2026).

\bibitem{Paper1b}
S.-K. Huang,
``Wavefunction Collapse as Retrodiction Failure: Gravitational
Decoherence, Objectivity, and the $\tau$-$R$ Bridge,''
companion paper (2026).

% === Petz recovery and quantum information ===

\bibitem{Petz1988}
D.~Petz,
``Sufficiency of channels over von Neumann algebras,''
Q.\ J.\ Math.\ \textbf{39}, 97 (1988).

\bibitem{Fawzi2015}
O.~Fawzi and R.~Renner,
``Quantum Conditional Mutual Information and Approximate
Markov Chains,''
Commun.\ Math.\ Phys.\ \textbf{340}, 575 (2015).

\bibitem{JRSWW2018}
M.~Junge, R.~Renner, D.~Sutter, M.~M.~Wilde, and A.~Winter,
``Universal Recovery Maps and Approximate Sufficiency of
Quantum Relative Entropy,''
Ann.\ Henri Poincar\'{e} \textbf{19}, 2955 (2018).

\bibitem{Parzygnat2023}
A.~J. Parzygnat and F.~Buscemi,
``Axioms for retrodiction: Achieving time-reversal symmetry
with a prior,''
Quantum \textbf{7}, 1013 (2023).

% === GR, refractive index, exponential metric ===

\bibitem{Einstein1911}
A.~Einstein,
``\"{U}ber den Einflu{\ss} der Schwerkraft auf die
Ausbreitung des Lichtes,''
Ann.\ Phys.\ (Leipzig) \textbf{35}, 898 (1911).

\bibitem{Dicke1957}
R.~H.~Dicke,
``Gravitation without a Principle of Equivalence,''
Rev.\ Mod.\ Phys.\ \textbf{29}, 363 (1957).

\bibitem{Hawking1973}
S.~W.~Hawking and G.~F.~R.~Ellis,
\emph{The Large Scale Structure of Space-Time}
(Cambridge University Press, 1973).

\bibitem{Papapetrou1954}
A.~Papapetrou,
``Eine Theorie des Gravitationsfeldes mit einer Feldfunktion,''
Z.\ Phys.\ \textbf{139}, 518 (1954).

\bibitem{Yilmaz1958}
H.~Yilmaz,
``New Approach to General Relativity,''
Phys.\ Rev.\ \textbf{111}, 1417 (1958).

\bibitem{Makukov2020}
M.~A.~Makukov and E.~G.~Mychelkin,
``Triple Path to the Exponential Metric,''
Found.\ Phys.\ \textbf{50}, 1346 (2020).

\bibitem{Boonserm2018}
P.~Boonserm, T.~Ngampitipan, A.~Simpson, and M.~Visser,
``Exponential metric represents a traversable wormhole,''
Phys.\ Rev.\ D \textbf{98}, 084048 (2018).

\bibitem{Misner1999}
C.~W.~Misner,
``Yilmaz Cancels Newton,''
Nuovo Cimento B \textbf{114}, 1079 (1999);
arXiv:gr-qc/9504050.

\bibitem{Mychelkin2024}
E.~G.~Mychelkin, M.~A.~Makukov, and G.~Suliyeva,
``On the weak and strong field effects in antiscalar
background,''
Gen.\ Relativ.\ Gravit.\ \textbf{56}, 44 (2024);
arXiv:2403.11610.

\bibitem{Bertotti2003}
B.~Bertotti, L.~Iess, and P.~Tortora,
``A test of general relativity using radio links with the
Cassini spacecraft,''
Nature \textbf{425}, 374 (2003).

\bibitem{ErnazarovIvashchuk2026}
K.~K.~Ernazarov and V.~D.~Ivashchuk,
``Phantom scalar field solution in the exponential metric,''
(2026).

\bibitem{NathSarma2025}
S.~Nath and D.~Sarma,
``Exponential wormhole solution in logarithmic $f(R)$ gravity,''
(2025).

\bibitem{Ernazarov2025}
K.~K.~Ernazarov,
``The Yilmaz-Rosen and Janis-Newman-Winicour metric solutions
in the scalar-Einstein-Gauss-Bonnet 4d gravitational model,''
arXiv:2510.21625 (2025).

\bibitem{Graber2017}
J.~R.~Graber,
``Rotating Yilmaz metric,''
arXiv:1708.05645 (2017).

\bibitem{MakukovMychelkin2023}
M.~A.~Makukov and E.~G.~Mychelkin,
``Rotating exponential metric via Newman--Janis algorithm,''
arXiv:2301.08118 (2023).

% === Gravitational decoherence and thermodynamics ===

\bibitem{Pikovski2015}
I.~Pikovski, M.~Zych, F.~Costa, and \v{C}.~Brukner,
``Universal decoherence due to gravitational time dilation,''
Nat.\ Phys.\ \textbf{11}, 668 (2015).

\bibitem{Tolman1930}
R.~C.~Tolman,
``On the Weight of Heat and Thermal Equilibrium in General
Relativity,''
Phys.\ Rev.\ \textbf{35}, 904 (1930).

\bibitem{LandiPaternostro2021}
G.~T.~Landi and M.~Paternostro,
``Irreversible entropy production: From classical to quantum,''
Rev.\ Mod.\ Phys.\ \textbf{93}, 035008 (2021).

\bibitem{Unruh1976}
W.~G.~Unruh,
``Notes on black-hole evaporation,''
Phys.\ Rev.\ D \textbf{14}, 870 (1976).

\bibitem{Bekenstein1981}
J.~D.~Bekenstein,
``Universal upper bound on the entropy-to-energy ratio for
bounded systems,''
Phys.\ Rev.\ D \textbf{23}, 287 (1981).

% === First-principles derivations of Sigma_grav ===

\bibitem{DorauMuch2025}
L.~Dorau and A.~Much,
``From Quantum Relative Entropy to the Semiclassical Einstein
Equations,''
Phys.\ Rev.\ Lett.\ (2025), arXiv:2510.24491.

\bibitem{Herrera2020}
L.~Herrera,
``Landauer Principle and General Relativity,''
Entropy \textbf{22}, 340 (2020).

\bibitem{AhmadiFuentes2014}
M.~Ahmadi, D.~E.~Bruschi, C.~Sabin, G.~Adesso, and I.~Fuentes,
``Relativistic Quantum Metrology,''
Sci.\ Rep.\ \textbf{4}, 4996 (2014).

% === Black hole information and holography ===

\bibitem{Susskind1993}
L.~Susskind, L.~Thorlacius, and J.~Uglum,
``The Stretched Horizon and Black Hole Complementarity,''
Phys.\ Rev.\ D \textbf{48}, 3743 (1993).

\bibitem{Page1993}
D.~N.~Page,
``Average entropy of a subsystem,''
Phys.\ Rev.\ Lett.\ \textbf{71}, 1291 (1993).

\bibitem{Penington2020}
G.~Penington,
``Entanglement Wedge Reconstruction and the Information Problem,''
J.\ High Energy Phys.\ \textbf{2020}(09), 002 (2020).

\bibitem{AEMM2021}
A.~Almheiri, T.~Hartman, J.~Maldacena, E.~Shaghoulian, and
A.~Tajdini,
``The entropy of Hawking radiation,''
Rev.\ Mod.\ Phys.\ \textbf{93}, 035002 (2021).

\bibitem{HaydenPreskill2007}
P.~Hayden and J.~Preskill,
``Black holes as mirrors: quantum information in random
subsystems,''
J.\ High Energy Phys.\ \textbf{2007}(09), 120 (2007).

\bibitem{Cotler2019}
J.~Cotler, P.~Hayden, G.~Penington, G.~Salton, B.~Swingle,
and M.~Walter,
``Entanglement Wedge Reconstruction via Universal Recovery
Channels,''
Phys.\ Rev.\ X \textbf{9}, 031011 (2019).

\bibitem{ADH2015}
A.~Almheiri, X.~Dong, and D.~Harlow,
``Bulk Locality and Quantum Error Correction in AdS/CFT,''
J.\ High Energy Phys.\ \textbf{2015}(04), 163 (2015).

\bibitem{CPS2020}
H.~Chen, G.~Penington, and G.~Salton,
``Entanglement Wedge Reconstruction using the Petz Map,''
J.\ High Energy Phys.\ \textbf{2020}(01), 168 (2020).

\bibitem{AMPS2013}
A.~Almheiri, D.~Marolf, J.~Polchinski, and J.~Sully,
``Black Holes: Complementarity vs.\ Firewalls,''
J.\ High Energy Phys.\ \textbf{2013}(02), 062 (2013).

% === Gravity from information ===

\bibitem{Jacobson1995}
T.~Jacobson,
``Thermodynamics of Spacetime: The Einstein Equation of State,''
Phys.\ Rev.\ Lett.\ \textbf{75}, 1260 (1995).

\bibitem{Bianconi2025}
G.~Bianconi,
``Gravity from entropy,''
Phys.\ Rev.\ D \textbf{111}, 066001 (2025);
arXiv:2408.14391.

\bibitem{Thattarampilly2026}
U.~Thattarampilly, Y.~Zheng, and V.~K.~Kakkat,
``Spherically symmetric black holes in Gravity from Entropy
and spontaneous emission,''
arXiv:2602.13694 (2026).

% === Horizons: apparent, quasi-local, QES ===

\bibitem{Hayward1994}
S.~A.~Hayward,
``General laws of black-hole dynamics,''
Phys.\ Rev.\ D \textbf{49}, 6467 (1994).

\bibitem{AshtekarKrishnan2004}
A.~Ashtekar and B.~Krishnan,
``Isolated and dynamical horizons and their applications,''
Living Rev.\ Relativ.\ \textbf{7}, 10 (2004).

\bibitem{Engelhardt2015}
N.~Engelhardt and A.~C.~Wall,
``Quantum extremal surfaces: holographic entanglement
entropy beyond the classical regime,''
J.\ High Energy Phys.\ \textbf{2015}(01), 073 (2015).

\bibitem{Kodama1980}
H.~Kodama,
``Conserved energy flux for the spherically symmetric system
and the backreaction problem in the black hole evaporation,''
Prog.\ Theor.\ Phys.\ \textbf{63}, 1217 (1980).

\bibitem{AbreuVisser2010}
G.~Abreu and M.~Visser,
``Kodama time: geometrically preferred foliations of GR
spacetimes,''
Phys.\ Rev.\ D \textbf{82}, 044027 (2010).

% === Observations ===

\bibitem{EHT2019}
Event Horizon Telescope Collaboration,
``First M87 Event Horizon Telescope Results.\ I.,''
Astrophys.\ J.\ Lett.\ \textbf{875}, L1 (2019).

\bibitem{Cardoso2016}
V.~Cardoso, E.~Franzato, and P.~Pani,
``Is the Gravitational-Wave Ringdown a Probe of the Event
Horizon?,''
Phys.\ Rev.\ Lett.\ \textbf{116}, 171101 (2016).

% === SYK and holography ===

\bibitem{SachdevYe1993}
S.~Sachdev and J.~Ye,
``Gapless spin-fluid ground state in a random quantum
Heisenberg magnet,''
Phys.\ Rev.\ Lett.\ \textbf{70}, 3339 (1993).

\bibitem{MaldacenaStanford2016}
J.~Maldacena and D.~Stanford,
``Remarks on the Sachdev-Ye-Kitaev model,''
Phys.\ Rev.\ D \textbf{94}, 106002 (2016).

\bibitem{GaoJafferisWall2017}
P.~Gao, D.~L.~Jafferis, and A.~C.~Wall,
``Traversable wormholes via a double trace deformation,''
J.\ High Energy Phys.\ \textbf{2017}(12), 151 (2017).

\bibitem{Jafferis2022}
D.~Jafferis, A.~Zlokapa, J.~D.~Lykken, D.~K.~Kolchmeyer,
S.~I.~Davis, N.~Lauk, H.~Neven, and M.~Spiropulu,
``Traversable wormhole dynamics on a quantum processor,''
Nature \textbf{612}, 51 (2022).

% === Recent related work ===

\bibitem{GalleyGiacominiSelby2023}
T.~D.~Galley, F.~Giacomini, and J.~O.~E.~Selby,
``Any consistent coupling between classical gravity and
quantum matter is fundamentally irreversible,''
Quantum \textbf{7}, 1142 (2023).

\bibitem{TrejoCalderon2025}
J.~A.~Trejo-Calder\'{o}n,
``Modular Channels, Thermal Filtering and the Spectral
Emergence of Spacetime,''
arXiv:2504.20457 (2025).

\bibitem{KasprzakTjoa2024}
M.~Kasprzak and E.~Tjoa,
``Transmission of quantum information through quantum fields
in curved spacetimes,''
arXiv:2408.00518 (2024).

\bibitem{Cirafici2024}
M.~Cirafici,
``Fluctuation theorems, quantum channels and gravitational
algebras,''
arXiv:2408.04219 (2024).

\bibitem{DorauVerch2024}
L.~Dorau and R.~Verch,
``Modular flow and entropy for axisymmetric spacetimes,''
arXiv:2402.18993 (2024).

\bibitem{Witten2022}
E.~Witten,
``Gravity and the crossed product,''
J.\ High Energy Phys.\ \textbf{2022}(10), 008 (2022).

\bibitem{Chandrasekaran2023}
V.~Chandrasekaran, R.~Longo, G.~Penington, and E.~Witten,
``An algebra of observables for de~Sitter space,''
J.\ High Energy Phys.\ \textbf{2023}(02), 082 (2023).

\bibitem{KudlerFlam2025}
J.~Kudler-Flam, S.~Leutheusser, and G.~Satishchandran,
``Generalized black hole entropy is von~Neumann entropy,''
arXiv:2309.15897 (2025).

\bibitem{supplement}
See Supplemental Material for detailed Kodama vector
derivation, event horizon calculations in Vaidya
spacetime, QES connection, and echo delay derivation.

\bibitem{Hiscock1981}
W.~A.~Hiscock,
``Models of evaporating black holes,''
Phys.\ Rev.\ D \textbf{23}, 2823 (1981).

\bibitem{BirrellDavies1982}
N.~D.~Birrell and P.~C.~W.~Davies,
\emph{Quantum Fields in Curved Space}
(Cambridge University Press, 1982).

\bibitem{IvanSabapathySimon2011}
J.~S.~Ivan, K.~K.~Sabapathy, and R.~Simon,
``Operator-sum representation for bosonic Gaussian channels,''
Phys.\ Rev.\ A \textbf{84}, 042311 (2011).

\bibitem{HolevoWerner2001}
A.~S.~Holevo and R.~F.~Werner,
``Evaluating capacities of bosonic Gaussian channels,''
Phys.\ Rev.\ A \textbf{63}, 032312 (2001).

\bibitem{Wald1994}
R.~M.~Wald,
\emph{Quantum Field Theory in Curved Spacetime and
Black Hole Thermodynamics}
(University of Chicago Press, 1994).

\end{thebibliography}

\end{document}
